---
title: "Apartment Analysis in Krakow"
author: "Yevhen Kuts"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Setting up my environment

Notes: setting up my R environment by loading the 'tidyverse', 'apartments_krakow_6month_final' and 'scales' packages

```{r ggplot visulisation for apartment krakow dataset}
library(ggplot2)
library(tidyverse)
library(scales)
```

```{r loading data}
krk_apartments <- read.csv("apartments_krakow_6month_final.csv")
# Display the first few rows to show a small understanding of the data
head(krk_apartments)
```

### Visualisations

Here we will go through a serious of visualisations

#### 1 Distribution of Apartment Prices in Krakow 2024

```{r}
ggplot(data=krk_apartments, aes(x = price)) +
  geom_histogram(binwidth = 10000, fill = "red", color = "white") +
  scale_x_continuous(
    breaks = seq(0, 2800000, by = 200000), # controling over tick positions
    labels = c("0", "200k", "400k", "600k", "800k",
               "1M", "1.2M", "1.4M", "1.6M", "1.8M",
               "2M", "2.2M", "2.4M", "2.6M", "2.8M") # controling labels
  ) +
  labs(
    title = "Distribution of Apartment Prices in Krakow",
    x = "Price (PLN)",
    y = "Number of Apartments"
  ) +   theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, face = "italic"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

#### 2 Apartments by Number of Rooms in Krakow

```{r}
ggplot(data=krk_apartments, aes(x = factor(rooms))) +
  geom_bar(fill = "purple", color = "black", width = 0.7) +
    geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = -0.5, # Position text above the bar
    color = "black",
    size = 2
  ) +
  labs(
    title = "Apartments by Number of Rooms in Krakow",
    x = "Number of Rooms",
    y = "Number of Apartments"
  ) +
  theme_minimal()
```

#### 3 Two-Variable Relationships 

``` {r}
ggplot(data = krk_apartments, aes(x = squareMeters, y = price)) +
  geom_point(alpha = 0.6, color = "darkblue") +
  geom_smooth(formula = y ~ x, method = "lm", color = "red", se = FALSE) +
  scale_y_continuous(
    breaks = seq(0, 3000000, by = 500000),
    labels = c("0", "500k", "1M", "1.5M", "2M",
               "2.5M", "3M")
  ) +
  labs(title = "Relationship Between Price and Apartment Size", x = "Square Meters", y = "Price (PLN)") +
  theme_minimal()
```

#### 4 Average Apartment Price by Number of Rooms

```{r}
ggplot(data=krk_apartments, aes(x = factor(rooms), y = price)) +
  stat_summary(fun = "mean", geom = "bar", fill = "gold", color = "blue") +
  labs(title = "Average Apartment Price by Number of Rooms", x = "Number of Rooms", y = "Average Price (PLN)") +
  theme_minimal()
```

### Conclusion

This analysis provides insights into apartment pricing trends, the influence of size and rooms on price, two-variable relationship and quick summary. Future work may include:

Analyzing data for other cities. Including temporal data to observe seasonal trends. Exploring additional features like balcony, own garden and differences by years.
